BlueTracer: \\ a Robust API Tracer for Evasive Malware

### 2018-07-06 implementation.tex
Tables Filled

We will start the current chapter by giving a general overview of BlueTracer. Specifically we will describe the main characteristics of Intel Pin, the DBI framework on which BlueTracer is based on. Then we will introduce BluePill, the DBI-based software toolkit allowing BlueTracer to be robust against anti-evasive malware. Lastly, we will illustrate BlueTracer's structure, which will serve as a baseline for the rest of the chapter.

@@

BlueTracer has been organized primarily taking into account the rich set of APIs offered by Pin, which have led to the decision to split the tool in three parts: the first aimed at \textbf{native APIs tracing}, the second for \textbf{APIs tracing} and the last focused on \textbf{callbacks and APCs tracing} (\textit{Figure 3.2}). In particular, the tracing of native APIs also employs a different source of API information (\textit{Dr. Memory}'s system call data) than the ones utilized for tracing APIs (\textit{drltrace}'s configuration file or the information extracted from \textit{PyREbox}'s database).

@@

>>>>

We will start the current chapter by giving a general overview of BlueTracer. Specifically we will describe the main characteristics of Intel Pin, the DBI framework on which BlueTracer is based on. Then we will introduce BluePill, the DBI-based software toolkit allowing BlueTracer to be robust against evasive malware. Lastly, we will illustrate BlueTracer's structure, which will serve as a baseline for the rest of the chapter.

@@

BlueTracer has been organized primarily taking into account the rich set of APIs offered by Pin, which have led to the decision to split the tool in three parts: the first aimed at \textbf{Native APIs tracing}, the second for \textbf{APIs tracing} and the last focused on \textbf{callbacks and APCs tracing} (\textit{Figure 3.2}). In particular, the tracing of native APIs also employs a different source of API information (\textit{Dr. Memory}'s system call data) than the ones utilized for tracing APIs (\textit{drltrace}'s configuration file or the information extracted from \textit{PyREbox}'s database).

@@

\section{Conclusions}

In this chapter we have provided a detailed description of how BlueTracer was organized and implemented. We have first introduced its building blocks, i.e., the Intel Pin DBI framework and the BluePill  anti-evasion software toolkit. We have then showed the tool's architecture, outlining its three core components: the Native API tracer, the API tracer and, the user-mode callbacks and APCs tracer. Next, we have explained how the issues related to multithreading were handled, and, finally, we have thoroughly analyzed each one of three aforementioned elements.

### 2018-07-06 introduction.tex
Tables Filled

A problem that all dynamic analysis techniques have to face, including function call monitoring by means of API hooking, is the widespread of evasive malware. The malware samples belonging to this category check whether or not they are being executed in an adverse environment and conceal their harmful behaviours accordingly, like for example by carrying out an exit sequence \cite{BluePill}. Unfortunately, such anti-evasion mechanisms are frequently adopted my malicious samples. According to Symantec's Internet Security Report of March 2018, 18\% of new malware were virtual-machine-aware \cite{Symantec}. To make matters worse, the API hooking techniques presented in literature are easily detectable and are not coupled by any mechanism to hide their presence from evasive malware.

\paragraph{Contributions.}
Throghout this thesis we present \textbf{BlueTracer}, a robust library and system call tracer for Windows applications, specialized in the monitoring of evasive malware. BlueTracer is based on the Intel Pin \cite{Pin} dynamic binary instrumentation (DBI) framework and is able to counteract malware's anti-evasion measures thanks to its integration with BluePill, a software toolkit built on top of a DBI layer which allows the simulation of the execution environment a particular malware sample was designed for and conceals any virtualization artifacts and setup details which might set off evasion \cite{BluePill}. BlueTracer is capable of tracing the input values, the output values and the return values of an extremely wide range of system calls (including Windows Native APIs) and API calls. Moreover, it also supports the tracing of Windows callbacks functions and Windows asynchronous procedure calls (APC). The tool was tested on a benign application aimed at assessing how good an anti-malware system is against evasion techniques and on actual evasive malware, proving to be effective in both tracing the samples' activity and remaining undetected.

>>>>

A problem that all dynamic analysis techniques have to face, including function call monitoring by means of API hooking, is the widespread of evasive malware. The malware samples belonging to this category check whether or not they are being executed in an adverse environment and conceal their harmful behaviours accordingly, like for example by carrying out an exit sequence \cite{BluePill}. Evasive techniques are frequently adopted my malicious samples. According to Symantec's Internet Security Report of March 2018, 18\% of new malware were virtual-machine-aware \cite{Symantec}. To make matters worse, the API hooking techniques presented in literature are easily detectable and are not coupled by any mechanism to hide their presence from evasive malware.

\paragraph{Contributions.}
Throughout this thesis we present \textbf{BlueTracer}, a robust library and system call tracer for Windows applications, specialized in the monitoring of evasive malware. BlueTracer is based on the Intel Pin \cite{Pin} dynamic binary instrumentation (DBI) framework and is able to counteract malware's evasion measures thanks to its integration with BluePill, a software toolkit built on top of a DBI layer which allows the simulation of the execution environment a particular malware sample was designed for and conceals any virtualization artifacts and setup details which might set off evasion \cite{BluePill}. BlueTracer is capable of tracing the input values, the output values and the return values of an extremely wide range of system calls (including Windows Native APIs) and API calls. Moreover, it also supports the tracing of Windows callbacks functions and Windows asynchronous procedure calls (APC). The tool was tested on a benign application aimed at assessing how good an anti-malware system is against evasion techniques and on actual evasive malware, proving to be effective in both tracing the samples' activity and remaining undetected.
